%
% SYDE 556/750 Assignments
%
% This is copyrighted material, you are not allowed to redistribute this file
% or parts thereof without permission from the copyright holders.
%
% (c) Chris Eliasmith
% (c) Andreas Stöckel
%

\input{../syde556_assignment_preamble.tex}

\newcommand{\assignmentCourse}{SYDE 556/750}
\newcommand{\assignmentCourseName}{Simulating Neurobiological Systems}
\newcommand{\assignmentName}{Assignment 3}
\newcommand{\assignmentNo}{03}
\newcommand{\assignmentTerm}{Winter 2020}
\newcommand{\assignmentDue}{March 3, 2020}
\newcommand{\assignmentMarks}{10 marks (10\% of the final grade)}
\newcommand{\assignmentURL}{https://github.com/astoeckel/syde556-w20/blob/master/assignments/assignment_03/syde556_assignment_03_template.ipynb}
\newcommand{\assignmentEmail}{astoecke@uwaterloo.ca}
\newcommand{\assignmentExtraNotes}{\item \textbf{Do not use or refer to any code from Nengo!}}

\begin{document}
	\assignmentHeader

	\section{Decoding from a population}
	
	As you did in previous assignments, make a population of $20$ LIF neurons representing a $1$-dimensional value, and compute a decoder for them. For parameters, $\tau_\mathrm{ref}=\SI{2}{\milli\second}$, $\tau_\mathrm{RC}=\SI{20}{\milli\second}$, the maximum firing rates are chosen randomly from a uniform distribution between $100$ and $\SI{200}{\hertz}$ (at the max radius), and the $x$-intercepts are chosen randomly from a uniform distribution between $-2$ and $2$. Remember that the $\alpha$ and $J^\mathrm{bias}$ terms are computed based on these $x$-intercepts and maximum firing rates.
		
	It is generally easiest to compute decoders using the original method from Assignment 1, where we use the rate-mode approximation for the neurons to generate the $A$ matrix, then find $\Gamma=A^T A + \sigma^2 I$. These decoders should work even when you simulate the neurons in terms of spikes (from question 2 on). However, you have to scale the output depending on the time step $\mathtt{dt}$.
	%The only difference will be that they will need to be scaled by $\Delta t$, your simulation time step.
		
	Use this method for computing decoders for the whole assignment.

	\begin{enumerate}[a)]
		\item \Marks{0.5} \emph{Tuning curves.} Plot the tuning curves (firing rate of each neuron for different $x$ values between $-2$ and $2$).
		\item \Marks{0.5} \emph{Decoder and error computation.} Compute the decoders and plot $(x-\hat{x})$. When computing decoders, take into account noise ($\sigma=0.1 \cdot \SI{200}{\hertz}$). When computing $\hat{x}$, add random Gaussian noise with $\sigma=0.1 \cdot \SI{200}{\hertz}$ to the activity. Report the Root Mean-Squared Error (RMSE).
	\end{enumerate}

	\section{Decoding from two spiking neurons}
	
	Choose a neuron from part 1 that has a firing rate of somewhere between \SIrange{20}{50}{\hertz} for $x=0$. Using that neuron's $\alpha$ and $J^\mathrm{bias}$ value, construct two neurons: both with the same $\alpha$ and $J^\mathrm{bias}$, but one with $e=1$ and the other with $e=-1$. With the function from the last assignment, generate a random input $x(t)$ that is \SI{1}{\second} long, with $\mathtt{rms}=1$, $\mathtt{dt}=\SI{1}{\milli\second}$, and an upper limit of \SI{5}{\hertz}. Feed that signal into the two neurons and generate spikes. Decode the spikes back into $\hat{x}(t)$ using a post-synaptic current filter $h(t)$ with a time constant of $\tau=\SI{5}{\milli\second}$.

	\begin{enumerate}[a)]
		\item \Marks{0.5} \emph{Synaptic filter.} Plot the post-synaptic current
		\begin{align*}
			h(t)= \begin{cases}
				0 & \text{if } t < 0 \,, \\
				\frac{e^{-t/\tau}}{\int_0^\infty e^{-t'/\tau} \mathrm{d}t'} & \text{otherwise} \,.
			\end{cases}
		\end{align*}
		\item \Marks{0.5} \emph{Decoding using a synaptic filter.} Plot the original signal $x(t)$, the spikes, and the decoded $\hat{x}(t)$ all on the same graph.
		\item \Marks{0.5} \emph{Error analysis.} Compute the RMSE of the decoding.
	\end{enumerate}

	\newpage
	
	\section{Decoding from many neurons}
	
	Repeat question 2, but with more neurons. Instead of picking particular neurons, randomly generate them with $x$-intercepts uniformly distributed between $-2$ and $2$ and with maximum firing rates between $100$ and \SI{200}{\hertz}. Randomly choose encoder values to be either $-1$ or $+1$.

	\begin{enumerate}[a)]
		\item \Marks{1.5} \emph{Exploring the error for an increasing neuron count.} Plot Root Mean-Squared Error as the number of neurons increases, on a log plot. Try $8$ neurons, $16$ neurons, $32$, $64$, $128$, up to $256$. For the RMSE for a particular number of neurons, average over at least $5$ randomly generated groups of neurons. For each group of neurons, randomly generate the signal $x(t)$. Use the same parameters as in question 2.
		\begin{itemize}
			\item[\symbolfont 🖈] The RMSE should go down as the number of neurons increases.
			\item[\symbolfont  📖] This should be similar to Figure 5.3 in the book.
		\end{itemize}
		\item \Marks{0.5} \emph{Discussion.} Discuss your results. What is the systematic relationship between the neuron count and the error?
	\end{enumerate}

	\section{Connecting two groups of neurons}
		
	For this question, use two groups of neurons with intercepts between $[-1, 1]$ to compute $y = 2x+1$. The first group of neurons will represent $x$ and the second group will represent $y$.
	
	Start by computing decoders. You will need two decoders: one to decode $f(x)=2x+1$ from the first population, and one to decode $f(y)=y$ (the standard representational decoder) from the second population. Remember that $\Upsilon$ can change depending on what function you want to decode.
	
	Use the same neuron parameters as for previous questions (other than the radius), and use $200$ randomly generated neurons in each population.

	\begin{enumerate}[a)]
		\item \Marks{0.5} \emph{Computing a function.} Show the behaviour of the system with an input of $x(t)=t-1$ for \SI{1}{\second} (a linear ramp from $-1$ to $0$). Plot the ideal $x(t)$ and $y(t)$ values, along with $\hat{y}(t)$.
		\begin{itemize}
			\item[\symbolfont 🖈] Note that you should use the decoders that work for any input over the range of intercepts: do not re-compute decoders for any particular input (i.e. set of $x$ values).
			Input $x(t)$ into the first group of neurons and produce spikes. Decode from those spikes using the decoder for $f(x)=2x+1$. Input that decoded result into the second group of neurons to produce spikes. Use the second decoder ($f(y)=y$) to decode $\hat{y}(t)$.
			\item[\symbolfont 🖈] Make sure the maximum firing rates are now at -1 or 1 (i.e., the radius is 1).
		\end{itemize}
		\item \Marks{0.5} \emph{Step input.} Repeat part (a) with an input that is ten randomly chosen values between -1 and 0, each one held for 0.1 seconds (a randomly varying step input)
		\item \Marks{0.5} \emph{Sinusoidal input.} Repeat part (a) with an input that is $x(t)=0.2\sin(6\pi t)$.
		\item \Marks{0.5} \emph{Discussion.} Briefly discuss the results for this question. Does the output match the ideal output? What kind of deviations do you see and why do those exist?
	\end{enumerate}

	\section{Connecting three groups of neurons}
		
	For this question, use three groups of neurons with intercepts from $[-1, 1]$ to compute $z = 2y+0.5x$. Follow the same steps as question 4, but take the decoded outputs from the first two groups of neurons ($f(y)=2y$ and $f(x)=0.5x$), add them together, and feed that into the third group of neurons.

	\begin{enumerate}[a)]
		\item \Marks{1} \emph{Sinusoidal input.} Plot $x(t)$, $y(t)$, the ideal $z(t)$, and the decoded $\hat{z}(t)$ for an input of $x(t)=\cos(3\pi t)$ and $y(t)=0.5 \sin (2 \pi t)$ (over \SI{1}{\second}).
		\item \Marks{0.5} \emph{Random input.} Plot $x(t)$, $y(t)$, the ideal $z(t)$, and the decoded $\hat{z}(t)$ for a random input over \SI{1}{\second}. For $x(t)$ use a random signal with a limit of \SI{8}{\hertz} and $\mathtt{rms}=1$. For $y(t)$ use a random signal with a limit of 5 Hz and $\mathtt{rms}=0.5$.
	\end{enumerate}


	\section{Computing with vectors}
	
	Do the same thing as questions 4 and 5, but with 2-dimensional vectors instead of scalars. Everything else is the same. For your encoders $e$, randomly generate them over the unit circle.
	
	The function to compute is $w = x-3y+2z-2q$. This requires five groups of neurons: $x$, $y$, $z$, $q$, and $w$. Each of them represents a $2$-dimensional value. The outputs from $x$, $y$, $z$, and $q$ all feed into $w$.
	
	\begin{enumerate}[a)]
		\item \Marks{1} \emph{Constant inputs.} Plot the decoded output $\hat{w}(t)$ and the ideal $w$ for
		\begin{align*}
			x &=(0.5,1), & y &= (0.1,0.3), & z &=(0.2,0.1), & q &= (0.4,-0.2) \,.
		\end{align*}
		\begin{itemize}
			\item[{\symbolfont 🖈}] Note that these are all constants, so they don't change over time. However, still plot the results for \SI{1}{\second} on one or more 2D graphs.
		\end{itemize}
		\item \Marks{0.5} \emph{Sinusoidal input.} Produce the same plot for
		\begin{align*}
			x &=(0.5,1), & y &= (\sin(4\pi t),0.3), & z &=(0.2,0.1), & q &= (\sin(4\pi t),-0.2) \,.
		\end{align*}
		\item \Marks{0.5} \emph{Discussion.} Describe your results and discuss why and how they stray from the expected answer.
	\end{enumerate}
\end{document}
