\documentclass[10pt,letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[svgnames]{xcolor}
\usepackage{graphicx}
\usepackage{csquotes}
\author{Andreas Stöckel}
\begin{document}
	
\newcommand{\Pred}[1]{\mathbf{\textcolor{Crimson}{#1}}}
\newcommand{\Obj}[1]{\mathtt{\textcolor{RoyalBlue}{#1}}}
\newcommand{\Fun}[1]{\mathit{\textcolor{ForestGreen}{#1}}}
\newcommand{\CC}{\circledast}

\paragraph{3.5 Revisiting Jackendoff's Challenges}
\begin{itemize}
	\item \textbf{The Binding Problem.} How to combine concepts into a new concept? $\Rightarrow$ Binding operators
	\begin{align*}	
		\Obj{RED} \CC \Obj{SQUARE} + \Obj{BLUE} \CC \Obj{CIRCLE}
	\end{align*}
	\item \textbf{The Problem of Two.} How can the same concept be active in two different contexts?
	$\Rightarrow $ Use symbols denoting the role. (Little star besides big star)
	\begin{align*}
		\Obj{OBJ1} \CC (\Obj{TYPE} \CC \Obj{STAR} + \Obj{SIZE} \CC \Obj{LITTLE}) + \\
		\Obj{OBJ2} \CC (\Obj{TYPE} \CC \Obj{STAR} + \Obj{SIZE} \CC \Obj{BIG}) + \\
		\Obj{REL} \CC \Obj{BESIDES}
	\end{align*}
	\item \textbf{The Problem of Variables.} How to express rules containing variables? $\Rightarrow$ Use inverse and role designators
	\begin{align*}
		\Obj{RULE} &= \Obj{RED} \CC \Obj{NOUN} \\
		\Obj{VAR} &= \Obj{NOUN}^{-1} \CC \Obj{BALL} \\
		\Obj{RULE} \CC \Obj{VAR} &\approx \Obj{RED} \CC \Obj{BALL}
	\end{align*}
	\item \textbf{Working vs. Long-Term Memory.}  How to manifest symbols in working and long-term memory?
	$\Rightarrow$ NEF: neural activities $\leadsto$ working memory, connection weights $\leadsto$ functions (e.g., associative memory)
\end{itemize}

\paragraph{THE SEMANTIC POINTER ARCHITECTURE}

\paragraph{1. Introduction}

\begin{itemize}
	\item VSAs: solve interesting cognitive tasks (Raven's Progressive Matrices, Jackendoff's Challenges)
	\item But: not a cognitive architecture $\rightarrow$ unclear how to build a behaving agent
	\item[$\Rightarrow$] Semantic Pointer Architecture (SPA)
	\begin{itemize}
		\item What is a semantic pointer?
		\item Working memory
		\item Action selection
	\end{itemize}
\end{itemize}

\paragraph{2. Semantic Pointers}

Combination of several concepts
\begin{itemize}
	\item VSAs (binding operator, circular convolution)
	\item NEF (computational substrate, neurons)
	\item Source of vectors: reversible compression/decompression (dereferencing $\leadsto$ semantic pointer, deep semantics)
	\item Architecture: basal ganglia/cortex control loop
	\item[$\Rightarrow$] can build small and large scale models
	\item[$\Rightarrow$] flexible architecture, solve multiple tasks within the same model (SPAUN)
\end{itemize}

\paragraph{2.1 Deep vs. Shallow semantics}
\begin{itemize}
	\item Related to the \enquote{symbol grounding problem}
	\item Shallow semantics: relational
	\item Deep semantics: symbols link at \enquote{subjective experience}
	\item Example for shallow semantics: random vectors, word embeddings
	\item word2vec example:
		\begin{align*}
		\Obj{WOMAN} + (\Obj{KING} - \Obj{MAN}) &\approx \Obj{QUEEN} \,,\\
		\Obj{SPACE} + (\Obj{SHIPS} - \Obj{WATER}) &\approx \Obj{SPACESHIP} \,.
		\end{align*}
\end{itemize}

\paragraph{2.2 Deep Semantics for Perception}
\begin{itemize}
	\item Example: visual processing in cortex
	\item Dimensionality reduction/compression from layer to layer
	\item First layer: high-dimensional visual input
	\item Final layer: semantic pointer
	\item Can decompress: clamp semantic pointer and compute most likely intermediate layer representations (i.e., \enquote{mental imagery})
\end{itemize}

\paragraph{2.3 Deep Semantics for Action}

\begin{itemize}
	\item Decompress low-dimensional semantic pointer
	\item Input: high-level action (\enquote{grab bottle})
	\item Output: body muscle tensions over time
	\item See HTBAB for an example (Chp. 3)
\end{itemize}

\paragraph{2.4 Using Semantic Pointers in Nengo}

\begin{itemize}
	\item \emph{See example}
\end{itemize}

\paragraph{2.5 SPAUN}

\begin{itemize}
	\item \emph{See slides}
\end{itemize}

\paragraph{3. Serial Working Memory}

\begin{itemize}
	\item Can use high-dimensional integrator for (working) memory
	\item Bind elements to position semantic pointer
	\begin{align*}
		\Obj{WM} &= \Obj{ITEM\_A} \CC \Obj{POS\_1} + \Obj{ITEM\_B} \CC \Obj{POS\_2} \\
		         &= \Obj{ITEM\_A} \CC \Obj{POS} + \Obj{ITEM\_B} \CC \Obj{POS} \CC \Obj{POS}
	\end{align*}
	\item In humans: primacy and recency effects
	\item Can model using two separate systems
\end{itemize}

\end{document}